Bootstrapping Multi-atlas Hippocampal Segmentation with \mb

### 2013-08-22 paper.Rnw
A few small changes based on Matt's feedback

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: 
%  - check tense: we are is have has 
%  - update montages with TAL coordinates
%  - add MAGeT brain diagram, and remove pseudocode
%  - check that \gs is used instead of Manual 
% Sweave/Knitr init
<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               fig.align = 'center', 
               tidy = FALSE, 
               comment = NA, 
               cache = TRUE,
               fig.width = 6, 
               fig.height = 6)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw(base_size = 12))

@@

To validate \mb segmentation quality with respect to other established automated
methods, \mb is applied to large dataset from the ADNI project and the resulting
segmentations are compared to those produced by FreeSurfer, FSL, MAPER and by
expert semi-automated manual segmentations (SNT).

\paragraph{ADNI1 dataset revisited} 
The {\em \adnidataset} standardized dataset contains $1919$ images in total.
SNT and FreeSurfer hippocampal volumes for a subset of images are provided by
ADNI, along with quality control data for each FreeSurfer segmentation
(guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
the entire \adnidataset dataset are shown in Table
\ref{tab:ADNI1-Dataset-Demographics}.

\paragraph{Experiment details} 
\mb was configured with an atlas library composed of the five Winterburn T1
atlases (described in Experiment 2), and the size of template library,
registration method, and label fusion method found in Experiment 1 were used.
The FSL FIRST segmentation method was used via the \verb+run_first_all+ script
according to the FIRST user guide
\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images in
the \adnidataset dataset were segmented by both methods.

One author (MP) performed visual quality inspection for MAGeT and FSL
segmentations using similar quality control guidelines (if either hippocampus
was under or over segmented by greater than 10mm in three or more slices then
the segmentation did not pass).  Only images meeting the conditions of having
segmentations from all methods (SNT, MAPER, FreeSurfer, FSL, and \mb) and also
passing quality control inspection are included in the analysis.

\paragraph{Evaluation method}
As in Experiment 3, the SNT \gs and Winterburn hippocampal segmentation
protocols differ in the neuroanatomical features delineated, and so we must
assess \mb by the degree of correlation of average hippocampal volume across all
subjects. Specifically, Pearson correlation is used. We also compute the
correlation in hippocampal volume between the existing, established automated
segmentation methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
volume-related fixed and proportional biases in all segmentation methods using
Bland-Altman plots \citep{Bland1986}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: 
%  - check tense: we are is have has 
%  - update montages with TAL coordinates
%  - add MAGeT brain diagram, and remove pseudocode
% Sweave/Knitr init
<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               fig.align = 'center', 
               tidy = FALSE, 
               comment = NA, 
               cache = TRUE,
               fig.width = 6, 
               fig.height = 6)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw(base_size = 12))

@@

To validate \mb segmentation quality with respect to other established
automated methods, \mb was applied to large dataset from the ADNI project and
the resulting segmentations were compared to those produced by FreeSurfer, FSL,
MAPER and the SNT \gs segmentations.

\paragraph{ADNI1 dataset revisited} 
The {\em \adnidataset} standardized dataset contains $1919$ images in total.
SNT and FreeSurfer hippocampal volumes for a subset of images were provided by
ADNI, along with quality control data for each FreeSurfer segmentation
(guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
the entire \adnidataset dataset are shown in Table \ref{tab:ADNI1-Dataset-Demographics}.

\paragraph{Experiment details} 
\mb was configured with an atlas library composed of the five Winterburn T1
atlases (described in Experiment 2), and size of template library,
registration method, and label fusion method were determined by the optimal
settings found in Experiment 1.  The FSL FIRST segmentation method was used via
the \verb+run_first_all+ script according to the FIRST user guide
\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
in the \adnidataset dataset were segmented by both methods.

One author (MP) performed visual quality inspection for MAGeT and FSL
segmentations using similar quality control guidelines (if either hippocampus
was under or over segmented by greater than 10mm in three or more slices then
the segmentation did not pass).  Only images meeting the conditions of having
segmentations from all methods (SNT, MAPER, FreeSurfer, FSL, and \mb) and also
passing quality control inspection were included in the analysis.

\paragraph{Evaluation method}
As in Experiment 3, the SNT \gs and Winterburn hippocampal segmentation
protocols differ in the neuroanatomical features delineated, and so we must
assess \mb by the degree of correlation of average hippocampal volume across all
subjects. Specifically, a Pearson correlation is used. We also compute the
correlation in hippocampal volume between the existing, established automated
segmentation methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
volume-related fixed and proportional biases in all segmentation methods using
Bland-Altman plots \citep{Bland1986}.
